Dynein project log

### 2016-12-16 thesis_stuff/thesis_stuff.tex
Add calculations for ct and free energy of binding from Redwine et. al.

\usepackage{marvosym} % Required for symbols in the colored box

@@

\begin{center}
  \begin{tabular}{| l | l | l | l | l | l |}
    \hline
    & Winch (Sarlah) & Schmidt & Lin & PyMol 3VKH & PyMol 4RH7\\\hline
    $L_s$ & 12nm & & & 21.0nm & 22.1nm\\ \hline
    $L_t$ &  7nm & & & & 11.15nm\\ \hline
    $R_b$ &  & & & 1.57nm & 1.45nm\\ \hline
    $R_m$ &  7nm & & & 7.36nm & 6.3nm\\ \hline
    $R_t$ &  & & & & 2.16nm \\ \hline
    $\theta_{m}^{Pre}$ & 250$^{\circ}$ & & 171$^{\circ}$ & &\\ \hline
    $\theta_{m}^{Post}$ & 330$^{\circ}$ & & 137.5$^{\circ}$ & &\\ \hline
    $\theta_{b}$ & 56$^{\circ}$ & & 63.5$^{\circ}$ & &\\ \hline
    $\theta_{t}$ & 0$^{\circ}$ & & & &\\ \hline
    $k_{ub}$ & $180 s^{-1,a}$ & & & &\\ \hline
    $k_b$ & $5000 s^{-1,b}$ & & & &\\ \hline
  \end{tabular}
\end{center}

@@

\subsection{Arrhenius Rate}
Papers often supply rate constants for various transitions. Our model uses Arhennius-type rates which take into account the change in conformational energy from switching equilibrium angles, so we must
convert between these two systems:

@@

\section{Interesting Crystal Structures}
\textbf{3VKH} 2.8A x-ray resolution full dynein motor, from Kon 2012 (2.8A crystal structure of the dynein motor domain). Probably what we'll use to get some equilibrium angles and lengths, though not sure if crystal structures are identical to in vivo structures. Discusses amino acid chemistry of the powerstroke and long-distance MTBD/ATPase communication.\\
\textbf{3J67/3J68} 34A EM structure of pre/post motor (from Structural Mechanism of dynein power stroke 2014). Shows in-depth changes to motor, but appears to only be the six ATPases and possibly buttress.
\textbf{3J1T/3J1U} 9.7A EM structure of tightly/loosly bound tubulin-MTBD binding from (Structural basis for microtubule binding and release by dynein).\\
\textbf{4RH7} 2.3A x-ray crystal structure of dynein motor/stalk with ADP.Vi (pre-stroke) from (Structure of human cytoplasmic dynein-2 primed for its power stroke). Could be another good place to get angles for pre-stroke, since 3VKH is post-stroke.\\

\section{Papers}
\subsection{Winch Model (dx.doi.org/10.1016/j.bpj.2014.06.022)}
-Communication between motors and BDs possible through $\alpha$-helix sliding. Bidirectional communication. [24,25]\\
-Pre, post-powerstroke states\\
-Pre-powerstroke: unbound state after ATP hydrolysis (our onebound)\\
-Post-powerstroke: bound state (our bothbound)\\
-Elastic stalk, unlike our rigid one\\
-Only tail angle changes between pre and post-powerstroke\\
-MTBD decides directionality?\\
-Tail-tail interaction, also motor-motor stacking interaction\\
-rate constant simulation, does not do dynamics to take into account time spent in a state\\

@@

>>>>

@@

\begin{center}
  \begin{tabular}{| l | l | l | l | l | l | l |}
    \hline
    & Winch (Sarlah) & Schmidt & Lin & PyMol 3VKH & PyMol 4RH7 & Redwine\\\hline
    $L_s$ & 12nm & & & 21.0nm & 22.1nm &\\ \hline
    $L_t$ &  7nm & & & & 11.15nm &\\ \hline
    $R_b$ &  & & & 1.57nm & 1.45nm &\\ \hline
    $R_m$ &  7nm & & & 7.36nm & 6.3nm &\\ \hline
    $R_t$ &  & & & & 2.16nm & \\ \hline
    $\theta_{m}^{Pre}$ & 250$^{\circ}$ & & 171$^{\circ}$ & & & \\ \hline
    $\theta_{m}^{Post}$ & 330$^{\circ}$ & & 137.5$^{\circ}$ & & &\\ \hline
    $\theta_{b}$ & 56$^{\circ}$ & & 63.5$^{\circ}$ & & &\\ \hline
    $\theta_{t}$ & 0$^{\circ}$ & & & & &\\ \hline
    $k_{ub}$ & $180 s^{-1,a}$ & & & & &\\ \hline
    $k_b$ & $5000 s^{-1,b}$ & & & & &\\ \hline
    $\Delta G_{bind, bb}$ & & & & & & -71 kJ/mol\\ \hline
    $\Delta G_{bind, ob}$ & & & & & & -37 kJ/mol\\ \hline
    $c_t$ & & & & & & \\ \hline
    $c_m$ & & & & & & \\ \hline
    $c_b$ & & & & & & 140 kJ/mol\\ \hline
  \end{tabular}
\end{center}

@@

\subsection{Finding Arrhenius preexpontential from k}
Papers often supply rate constants for various transitions. Our model uses Arhennius-type rates which take into account the change in conformational energy from switching equilibrium angles, so we must convert between free energy and rates:

@@

\subsection{Calculating $\Delta G_{bind}$}
Based on Van der Spoel \textit{et. al.}, $\Delta H_{HB} = 10.6kJ/mol$ for water molecules forming a single hydrogen bond. Based on another paper (Anderson \textit{et al}), the
``stabilizing energy'' of a single salt bridge (H-bond between charged molecules) is about 16 kJ/mol.\\

Based on the interpreted bonds from crystal structures of the MTBD-MT complex shown in Redwine \textit{et. al}, there are 5.9 salt bridges in the high-affinity MTBD-MT
complex and 3.1 in the low-affinity complex. Thus if we guess the free energy of a salt bridge as 12 kJ/mol, we can estimate binding energies as:

\begin{align*}
  \Delta G_{bind, high-affinity} &= -71 \mbox{ kJ/mol}\\
  \Delta G_{bind, low-affinity} &= -37 \mbox{ kJ/mol}\\
\end{align*}

\subsection{Estimating $c_b$ spring constant}
If the energy of binding of the MTBD foot to the MT is -71 kJ/mol, we can get a rough estimate of the spring constant for the binding domain. We say that
$\Delta \theta \approx 0$ corresponds to being fully bound and $\Delta \theta \approx 1$ corresponds to breaking all salt bridges and being unbound. Thus:

\begin{align*}
  \frac 12 c_b \Delta \theta_{ba} = \Delta G_{bind, high} &\rightarrow c_b = 2\Delta G_{bind, high} \approx 140 \mbox{ kJ/mol}\\
\end{align*}

\section{Papers}

\subsection{Interesting Crystal Structures}
\textbf{3VKH} 2.8A x-ray resolution full dynein motor, from Kon 2012 (2.8A crystal structure of the dynein motor domain). Probably what we'll use to get some equilibrium angles and lengths, though not sure if crystal structures are identical to in vivo structures. Discusses amino acid chemistry of the powerstroke and long-distance MTBD/ATPase communication.\\
\textbf{3J67/3J68} 34A EM structure of pre/post motor (from Structural Mechanism of dynein power stroke 2014). Shows in-depth changes to motor, but appears to only be the six ATPases and possibly buttress.
\textbf{3J1T/3J1U} 9.7A EM structure of tightly/loosly bound tubulin-MTBD binding from (Structural basis for microtubule binding and release by dynein).\\
\textbf{4RH7} 2.3A x-ray crystal structure of dynein motor/stalk with ADP.Vi (pre-stroke) from (Structure of human cytoplasmic dynein-2 primed for its power stroke). Could be another good place to get angles for pre-stroke, since 3VKH is post-stroke.\\

\subsection{H-bond $\Delta G$ calculation paper}
van der Spoel, D., J. van Maaren, P., Larsson, P., Timneanu, N. ``Thermodynamics of Hydrogen Bonding in Hydrophilic and Hydrophobic Media'' \textit{J. Phys. Chem. B} 2006: 110, 4393-4398.\\

This paper uses classical simulations of water and alcohols to calculate thermodynamic parameters for the transition state and equilibrium state of water-water and water-alcohol
hydrogen bonds. It uses H-bond lifetime and the Eyring rate equation to calculate transition state $\Delta G$ and number of H bond pairs / unbonded pairs to find equilibrium
$\Delta G$. It then uses a Van't Hoff calculation to find enthalpies and entropies. Raman spectroscopy is shown to find a different enthalpy for water-water H-bonds. Findings:

\begin{center}
  \begin{tabular}{| l | l | l | l | l | l |}
    \hline
    & $\Delta G$ & $\Delta H$ & $T\Delta S$ \\\hline
    Simulation & 5.11 kJ/mol & 9.0 kJ/mol & 4.1 kJ/mol\\ \hline
    Experimental & 10.6 kJ/mol & & \\ \hline
  \end{tabular}
\end{center}

where T = 298.15K.\\

Another study (pH-Induced denaturation of proteins: a single salt bridge contributes 3-5 kcal/mol to the free energy of folding of T4 lysozyme by D. Anderson) found that
a single salt bridge contributes about 16 kJ/mol to stabilizing a protein. We can use these numbers as estimates of the binding energy of the motor.\\

\subsection{MTBD-Tubulin complex}
Redwine, W.B. \textit{et. al.} ``Structural basis for microtubule binding and release by dynein. '' Science 2012, 337(6101): 1532–1536.

This is a sub-nanometer MTBD-tubulin complex structure which shows how the complex forms. A cryo-EM structure was acquired, atomic-resolution structures for the high-affinity
MTBD and tubulin dimers were superimposed.  Rigid body docking and molecular dynamics was then used to optimize this structure. Another technique was used to figure out how
the low-affinity MTBD would bind, and yet another technique was used to translate between the low and high-affinity MTBDs to make an animation. Also did this with the Kon
vanadium onebound crystal structure and not much changed.\\

The MTBD-tubulin dimers are on the PDB as 3J1T for the high-affinity and 3J1U for the low-affinity structures.\\

After finding the high- and low-affinity structures for the complexes, the authors guess at what salt bridges form in the two states between MT and MTBD. The reported values are shown in the tables below.\\

\begin{table}
  \parbox{.45\linewidth}{
    \centering
    \begin{tabular}{| l | l |}
      \hline
      Residue & Occupancy \\\hline
      R3382-αE417 & 23\% \\\hline
      R3382-αE420 & 79\% \\\hline
      R3382-αG416 & 40\% \\\hline
      K3299-βE431 & 60\% \\\hline
      K3299-βD427 & 53\% \\\hline
      R3342-βE159 & 51\% \\\hline
      N3310-αG410 & 38\% \\\hline
      L3385-αE415 & 48\% \\\hline
      S3384-αG415 & 27\% \\\hline
      A3383-αE415 & 49\% \\\hline
      R3306-βE196 & 86\%*\\\hline
      R3337-βD199 & 2\%* \\\hline
      R3342-βN197 & 36\%*\\\hline
    \end{tabular}
    \caption{High-affinity MTBD-MT complex salt bridges. RH column is bonds between MTBD and MT, respectively.}
  }
  \hfill
  \parbox{.45\linewidth}{
    \centering
    \begin{tabular}{| l | l |}
      \hline
      Residue & Occupancy \\\hline
      A3383-αE415  & 45\% \\ \hline
      R3382-αG416  & 44\% \\ \hline
      R3382-αE417  & 61\% \\ \hline
      R3382-αE420  & 79\% \\ \hline
      K3299-βE196  & 27\% \\ \hline
      S3307-βR253  & 16\% \\ \hline
      L3385-αE415  & 34\% \\ \hline
      N3310-αK112  & 10\% \\ \hline
    \end{tabular}
    \caption{Low-affinity MTBD-MT complex salt bridges. RH column is bonds between MTBD and MT, respectively.}
  }
\end{table}

From this table we count the number of effective salt bridges in the high/low affinity complexes: 5.9 and 3.1, respectively.\\

Some \textbf{really} cool videos of the MTBD switching from high/low affinity states. If you could get a similar video of the AAA domains binding ATP, these would be really
cool videos to put in a presentation.\\

\subsection{Winch Model (dx.doi.org/10.1016/j.bpj.2014.06.022)}
-Communication between motors and BDs possible through $\alpha$-helix sliding. Bidirectional communication. [24,25]\\
-Pre, post-powerstroke states\\
-Pre-powerstroke: unbound state after ATP hydrolysis (our onebound)\\
-Post-powerstroke: bound state (our bothbound)\\
-Elastic stalk, unlike our rigid one\\
-Only tail angle changes between pre and post-powerstroke\\
-MTBD decides directionality?\\
-Tail-tail interaction, also motor-motor stacking interaction\\
-rate constant simulation, does not do dynamics to take into account time spent in a state\\

@@

\subsection{Further reading}
\begin{itemize}
\item The dynein family at a glance. Höök P, Vallee RB J Cell Sci. 2006 Nov 1; 119(Pt 21):4369-71. The actual differences between axonemal/cytoplasmic dynein.
\end{itemize}
